---
title: "Wine About It Why Doncha"
author: "Evan"
date: "01/03/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# set root to rproj root
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(tidyverse)
library(readxl)
library(plotly)
```


## The Quest for the Best

Using data acquired during a night of garlic breath and shots of wine, we were tasked with finding the preferred red and white wine.

Below is a glimpse of the dataset.

```{r cars}
# read the data
wines_df  <- read_xlsx("data/garlic_games_wine_tasting.xlsx", sheet = "wine_tasting")
sample_n(wines_df, 10) %>% 
        knitr::kable()
```

## Boxplots of Ratings

The spread of ratings are shown below, using a boxplot to show the range of ratings for each wine.  Hover values show individual scoring, along with comments.

```{r echo=FALSE}


p <- wines_df %>% 
        ggplot(aes(x = fct_reorder(wine_name, colour), y = rating, text = comments, fill = colour)) + 
        geom_boxplot() + 
        geom_point() + 
        scale_y_continuous(limits = c(0, 20)) + 
        scale_fill_manual(values = c("#bf180a", "#ccc6c6")) + 
        theme_bw() + 
        theme(axis.text.x = element_text(angle = 30, hjust = 1)) + 
        coord_flip() + 
        labs(y = "Rating out of 20", x = "Wine Name", title = "Distribution of Wine Ratings by Type")

ggplotly(p)
```


## Boxplots of People

Do people tend to vote a certain way?

```{r echo=FALSE}


p <- wines_df %>% 
        ggplot(aes(x = wine_tester, y = rating, text = comments)) + 
        geom_boxplot() + 
        geom_point() + 
        scale_y_continuous(limits = c(0, 20)) + 
        scale_fill_manual(values = c("#bf180a", "#ccc6c6")) + 
        theme_bw() + 
        theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
        labs(y = "Rating out of 20", x = "Person", title = "Distribution of Wine Ratings by Person") 

ggplotly(p)
```

How about by colour by person?

```{r echo=FALSE}


p <- wines_df %>% 
        ggplot(aes(x = wine_tester, y = rating, text = comments, fill = colour)) + 
        geom_boxplot() + 
        geom_point() + 
        scale_y_continuous(limits = c(0, 20)) + 
        scale_fill_manual(values = c("#bf180a", "#ccc6c6")) + 
        theme_bw() + 
        theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
        labs(y = "Rating out of 20", x = "Person", title = "Distribution of Wine Ratings by Person") +
        facet_grid(.~colour)

ggplotly(p)
```

## Regression

Estimation of wine and person effect on score - try two different models:

1) Only wines: Y = X0 + B1X1
   - Y = rating
   - X0 = intercept
   - B1 = effect of wine type on rating (a slope)
   - X1 = wine types
2) Only people: Y = X0 + B2X2
   - B2 = effect of person on rating (a slope)
   - X2 = people


### Linear Model #1

```{r}
wines_only_lm <- lm(rating ~ wine_name, data = wines_df)

summary(wines_only_lm)
```

Black Shook Shiraz and San Pedro 1865 Cab Sav are the only ones with a positive slope; note that everything is compared to the reference group, 14 Hands Merlot.  Also under the p-value significance assumption of 0.05, no findings are sigificant.  This is to be expected from a small sample size such as this one.

### Linear Model #2

```{r}
wines_people_lm <- lm(rating ~  wine_tester, data = wines_df)

summary(wines_people_lm)
```

This one is quite funny - Adam is the reference group for people so all Estimates are in reference to him.  As everyone has a negative coefficient (Estimate column), we are all wine snobs compared to Adam.

This time there are a few significant results (noted by the * or **).  It seems that Carly and Kait are statistically significant wine snobs when compared to Adam.


## Conclusion